---
title: 'CA 3'
author: 'Daragh Brady'
date: '2024-02-20'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## CA 2 - D00255640

(a) Describe the data, referring to the number of variables and sample size, type of variables present in the data.

```{r}
path='C:/Users/darag/OneDrive/Documents/Statistics Labs/CA 2/car_prices.csv'
data = read.csv(path)
```

```{r}
ncol(data)
nrow(data)
str(data)
```

(b) Make suitable visualisations and tables to explore the data and interpret the ones you have selected for your report to tell the story about the data.

```{r}
hist(data$Year, main = 'Histogram of Year', xlab = 'Year')
plot(data$Price, main = 'Scatterplot of Price')
pie(table(data$Transmission), main = 'Pie Chart of Transmission')
pie(table(data$Fuel_Type), main = 'Pie Chart of Fuel type')
summary(data$Year)
summary(data$Price)
table(data$Fuel_Type, data$Transmission)
```

From executing this cell, I can see that from the histogram, the data has the most records around the year 2015. The price of cars are quite low, as the scatterplot shows that throughout the data the vast majority of dots are between 0 and 50, which means 0 and 5000000 INR Lakh, and through doing the summary function on Price, I found that the mean value is 940500 INR Lakh. Through using pie charts, I found that the majority of the Transmission column is Manual cars, and the Fuel type of the data is almost a half and half split.
