---
title: "CA 3"
author: "Daragh Brady"
date: "2024-02-20"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## CA 2 - D00255640

**a. Describe the data, referring to the number of variables and sample size, type of variables present in the data.**

```{r}
path='C:/Users/darag/OneDrive/Documents/Statistics Labs/CA 2/car_prices.csv'
data = read.csv(path)
```

```{r}
ncol(data)
nrow(data)
str(data)
```

After performing this code, I found the basic information about the csv file. I found the number of columns was 11, and the number of rows is 5532. I also found the type of variables in this csv file.

**b. Make suitable visualisations and tables to explore the data and interpret the ones you have selected for your report to tell the story about the data.**

```{r}
par(mfrow=c(2,2))
hist(data$Year, main = 'Histogram of Year', xlab = 'Year')
plot(data$Price, main = 'Scatterplot of Price')
pie(table(data$Transmission), main = 'Pie Chart of Transmission')
pie(table(data$Fuel_Type), main = 'Pie Chart of Fuel type')
summary(data$Year)
summary(data$Price)
table(data$Fuel_Type, data$Transmission)
```

From executing this cell, I can see that from the histogram, the data has the most records around the year 2015. The price of cars are quite low, as the scatterplot shows that throughout the data the vast majority of dots are between 0 and 50, which means 0 and 5000000 INR Lakh, and through doing the summary function on Price, I found that the mean value is 940500 INR Lakh. Through using pie charts, I found that the majority of the Transmission column is Manual cars, and the Fuel type of the data is almost a half and half split. After using the summary function on the year column, I found the maximum value is 2019 and minimum is 1998, this shows there is some outliers as the first quantile is 2012 which is closer to the maximum value (7) than the minimum value(14).

**c. Discuss potential research questions that can be analysed using this data, what the main response variable is in the data set. Is there anything to note about this response variable.**

From analysing the dataset, I found the response variable is Price. This leads to some possible research questions to be, How does Price vary with the factors: Fuel type, Transmission, Owner type? Is there a relationship between the Kilometres driven and Price? Is there a relationship with Engine size, Power and Price? How does the Year of car affect the Price? The main things to notes about the response variable is the outliers, for extremely cheap and expensive cars, and the context of the price, as certain brands, features and condition can be impacting factors on the Price.

**d. Construct an appropriate model to predict the response variable from part c. Make sure the assumptions are met and if they are not met can you make any adjustments to modelling approach to satisfy the assumptions. Once you are happy with the modelling assumptions being met interpret your model.**

```{r}
original_model <- lm(Price ~ Kilometers_Driven + Fuel_Type + Transmission + Owner_Type + Engine + Power + Seats + Year, data = data)

summary(original_model)
par(mfrow=c(2,2))
plot(original_model)

```

From this code block, I find that there is not much of a relationship between all of the affecting factors and Price, as the r squared value of 0.6795 shows the lack of linearity in this selected data. It is also proven through the residuals vs fitted plot and scale location model as the dots do not match the expected line, this shows there is not much homoscedasticity. Also shown in the Q-Q residuals plot the line does not fit the model well, as it tails off at the start and end of the expected line. The residuals vs levereage plot also is not expected as the majority of dots are at the side closest 0 and off the dashed line. The Cook's distance model also does not fit the model either as it shows that the model has a very large influence. Overall, this proves that this model is not good at predicting the Price variable with these factors.

**e. Are you happy with the fit of the model from part d? Is there anything you would change? Justify your answer.**

After reviewing the previous code block, I found this model to be not very good, as the relationship between the Price and the other numerical and categorical factors proved to be not very accurate. The graphs generated from the model are also not very good, as they do not meet the expected fittedness of the different graphs generated. I would remove the factors that have very little influence on predicting the Price. I would remove the Kilometers_Driven variable as it shows little impact on the model.

**f. Make the adjustments specified in e and fit this new model. Are the assumptions met? If the assumptions are met, interpret the results for the regression.**

```{r}
library(car)
new_model <- lm(log(Price) ~ log(Kilometers_Driven) + Fuel_Type + Transmission + Power + Year, data = data)
summary(new_model)

vif(new_model)
par(mfrow=c(2,2))
plot(new_model)
```

This model is a lot better, the changes that I have made have made the model more useful. The factors that I have included in this model are all of the highest significance so they should be left in. I have also used the log of Price and Kilometres Driven as it makes the model improve linearity, and it makes the variance more even. The plots are a lot better as they match the expected output of each graph. The VIF values indicate that there is moderate correlation with other predictors.

**g. Can the models run in part d and f be compared? If so compare and Conclude which model is preferred.**

```{r}
cat('Adjusted R-squared for new_model:',summary(new_model)$adj.r.squared,'\n')
cat('Adjusted R-squared for original_model:',summary(original_model)$adj.r.squared,'\n')

cat('AIC for new_model:',AIC(new_model),'\n')
cat('AIC for original_model:',AIC(original_model),'\n')

cat('BIC for new_model:',BIC(new_model),'\n')
cat('BIC for original_model:',BIC(original_model),'\n')

par(mfrow=c(2, 2))
plot(new_model)
plot(original_model)
```

From executing this code block, I found that the new model's adjusted R squared value of 0.865592 is higher than the original models' value of 0.6789677 showing that the new model explains the response variable's variability better than the original model. The AIC value for the new model is 2654.128 compared to 35611.91 from the original model, this means that the new model is a far better fit for the data. The BIC value of 2700.456 for the new model indicates that it is a better fit than the 35684.72 value of the original model, as it is considerably lower.
